---
title: "M215_Proj_Draft"
output: html_document
---

#### Name:Ziyan Song, Ke Zhao, Huiyu Hu

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(survival)
```

1. Read in data
```{r}
col.name <- c('survival', 'alive', 'ha.age', 'pe', 'fs', 'epss', 'lvdd', 'wm.score', 'wm.index', 
              'mult', 'name', 'group', 'alive.1yr')
col.type <- cols(col_double(), col_integer(), col_double(), col_integer(), col_double(), 
                 col_double(), col_double(), col_double(), 
                 col_double(), col_double(), col_character(), col_integer(), col_integer())
# Raw data
ec.data <- read_delim('/Users/huiyuhu/Desktop/Study/UCLA_Biostat/M215/M215_Project/echocardiogram.data', 
                      delim = ',', col_names = col.name, col_types = col.type)

# delete useless columns
#ec.data <- ec.data %>% select(-one_of('name', 'group', 'mult'))
ec.data <- ec.data[,-c(6,8,10,11,12,13)]
ec.data <- ec.data %>% mutate(death = ifelse(alive == 1, 0, 1))
# delete pc incorrect data
ec.data <- ec.data %>% filter(pe == 0 | pe == 1)
head(ec.data)

```

2. Statistic analysis

#### Censored data and missing data
```{r}
# percentage of the censored data
ec.data %>% group_by(alive) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n))
# censored data: 32.6%, so this dataset is good to use

# count how many missing for variables in each alive group
ec.data %>% group_by(alive) %>% summarise_all(funs(sum(is.na(.))))

ec.data %>% group_by(alive) %>% summarise(mean_age = mean(survival, na.rm = T))
```

#### Correlation

```{r}
# Data Exploration and Visualization
cor(ec.data, use="pairwise.complete.obs")
```

#### Data Exploration and Visualization

* Age
```{r}
# Age histogram
ec.data %>% filter(!is.na(ha.age)) %>%
  ggplot(aes(x=ha.age)) + 
  geom_histogram(color="darkblue", fill="lightblue") + 
  labs(title="Age histogram plot",x="Age (yr)", y = "N") + 
  theme_classic()
# Almost all patients are older than  40. Majority are from 55 to 70


```

* Fractional-shortening

```{r}
# fractional-shortening
ec.data %>% filter(!is.na(fs)) %>%
  ggplot(aes(x=fs)) + 
  geom_histogram(color="darkgreen", fill="lightgreen") + 
  labs(title="FS histogram plot",x='Fractional shortening', y = "N") + 
  theme_classic()
# Normal range is 25-45%, Mild is 20-25%, Moderate is 15-20%, and Severe is <15%
# Convert the FS to categorical variable 1:normal, 0:abnormal
ec.data <- ec.data %>% mutate(fs.cat = ifelse(fs > 0.25, 1, 0))

ec.data <- ec.data %>% mutate(fs.cat.2 = case_when(fs <= 0.25  ~ 'Abnormal',
                                      fs > 0.25 ~ 'Nomal'))

```

* Left ventricular wall motion score index: The shortening fraction reflects the left ventricle systolic function

```{r}
# Left ventricular wall motion score index (WMSI), defined as the average wall motion score of 17 myocardial segments (1 = normal, 2 = hypokinetic, 3 = akinetic, 4 = dyskinetic scored for each segment) was assessed at all 3 stages of stress.

ec.data %>% filter(!is.na(wm.index)) %>%
  ggplot(aes(x=wm.index)) + 
  geom_histogram(color="orange", fill="yellow") + 
  labs(title="WMSI histogram plot",x='WMSI', y = "N") + 
  theme_classic()

ec.data <- ec.data %>% mutate(wmsi = ifelse(wm.index <= 1.25, 1, 0))
# 1: normal, 0: hypokinetic, akinetic


```

* left ventricular end-diastolic dimension. This is a measure of the size of the heart at end-diastole.

```{r}
ec.data %>% filter(!is.na(lvdd)) %>%
  ggplot(aes(x=lvdd)) + 
  geom_histogram(color="darkred", fill="lightpink") + 
  labs(title="LVDD histogram plot",x='LVDD (in cm)', y = "N") + 
  theme_classic()

ec.data <- ec.data %>% mutate(lvdd.cat = ifelse(lvdd<= 5.0, 1, 0))




```


3. Fit model 

### a) Non-parametric methods

#### Kaplan-Meier estimator 

* For survival data one commonly wants to find an estimate of the survival function S(t). The standard estimator of S(t) is the Kaplan-Meier estimator.  Kaplan-Meier estimator is nonparametric, which requires no parametric assumptions. 

```{r}
fit.km <- survfit(Surv(survival, death) ~ 1, data = ec.data)
summary(fit.km)

# Plotting the Kaplan-Meier Estimates
plot(fit.km, main = 'Product-Limit Survival Estimates',
ylab = 'KM Estimator without Covariates', xlab = 'Time (in Month)')

# Plotting -log[S(t)] vs. t
plot(-log(fit.km$surv) ~ fit.km$time, type = 's', col = 'blue', 
     main = 'Estimated Cumulative Hazard Function', ylab = 'H(t) = -log[S(t)]',
     xlab = 'Time (in Month)')


```
* The median is **29**.

#### Nelson-Aalen Estimator

```{r}
H.km <- -log(fit.km$surv) 
#Nelson-Aalen Estimate of H(t):
h <- fit.km$n.event/fit.km$n.risk
H.na <- cumsum(h) #Nelson-Aalen estimates
s <- fit.km$n.event/fit.km$n.risk^2
V.na <- cumsum(s) #Variance estimate for the N-A estimator
NAest <- cbind(H.na, sqrt(V.na))
colnames(NAest) <- c("NA-Est.","Std. Err")

#Plot of the Nelson-Aalen Estimate
plot(H.na ~ fit.km$time, type = 's', 
     main="Plot of Nelson-Aalen Estimates", xlab="Time (t)",
     ylab="Estimated Cum. Hazard Function", col = "red")
lines(-log(fit.km$surv) ~ fit.km$time, type = "s", lty = 2, col = "blue")
legend("topleft", c( "Nelson-Aalen","Kaplan-Meier"), lty = 1:2,
col = c("red", "blue"), bty = "n", cex = .9)

# plot(-log(fit.km$surv) ~ fit.km$time, type = 's', col = 'blue', 
#      main = 'Estimated Cumulative Hazard Function', ylab = 'H(t) = -log[S(t)]',
#      xlab = 'Time (in Month)')
```

#### Comparing the Nelson-Aalen Estimate to the Kaplan-Meier Estimate
```{r}
plot(H.km~fit.km$time, xlab = "time", ylab = " Estimated Cumulative Hazard Function", 
     main = "Different H(t) Estimates (Non-parametric)", type = "s", col = "red")
lines(H.na~fit.km$time, type = "s", lty = 2, col = "blue")
legend("bottomright", c("Kaplan-Meier", "Nelson-Aalen"), lty = 1:2,
col = c("red", "blue"), bty = "n", cex = .9)
```

#### Stratified Analyses

- Create stratified Kaplan-Meier curves by groups

- Stratify on pericardial-effusion

```{r}
# pericardial-effusion
# summarize the frequency
ec.data %>% group_by(pe) %>%
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n))
# Only 18.3% of patient whose pericardial effusion is fluid around the heart

fit.km1 <- survfit(Surv(survival, death) ~ pe, data = ec.data)
plot(fit.km1, col = c('red', 'blue'), lty = c(1, 2), ylab = 'Estimated Disease-Free Survival', 
     xlab = 'Time', main = 'Kaplan-Meier Estimates by pericardial-effusion')
legend("topright", c("no fluid", "fluid"), lty = 1:2, 
       col = c("red", "blue"), bty = "n", cex = .9)

# Log-rank test
#Log-Rank Test (W = 1)
survdiff(Surv(survival, death) ~ pe, data = ec.data, rho = 0)

survdiff(Surv(survival, death) ~ fs.cat + wmsi + strata(pe), data = ec.data)
```

- Stratify on fractional-shortening

```{r}
ec.data %>% group_by(fs.cat) %>%
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n))
# Normal range is 25-45%
fit.km1 <- survfit(Surv(survival, death) ~ fs.cat, data = ec.data)
plot(fit.km1, col = c('red', 'blue'), lty = c(1, 2), 
     ylab = 'Estimated Disease-Free Survival', xlab = 'Time', 
     main = 'Kaplan-Meier Estimates by Fractional-Shortening')
legend("topright", c("Abnormal (less than 25%)", "Normal (25-45%)"), lty = 1:2, 
       col = c("red", "blue"), bty = "n", cex = .9)

#Log-Rank Test (W = 1)
survdiff(Surv(survival, death) ~ fs.cat, data = ec.data, rho = 0)

survdiff(Surv(survival, death) ~ strata(fs.cat) + wmsi + pe, data = ec.data)

```

- Stratify on wmsi

```{r}
ec.data %>% group_by(wmsi) %>%
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n))
# 1: normal, 0: hypokinetic, akinetic

fit.km1 <- survfit(Surv(survival, death) ~ wmsi, data = ec.data)
plot(fit.km1, col = c('red', 'blue'), lty = c(1, 2), 
     ylab = 'Estimated Disease-Free Survival', xlab = 'Time', 
     main = 'Kaplan-Meier Estimates by WMSI')
legend("topright", c("Hypokinetic & Akinetic", "Normal"), lty = 1:2, 
       col = c("red", "blue"), bty = "n", cex = .9)
#Log-Rank Test (W = 1)
survdiff(Surv(survival, death) ~ wmsi, data = ec.data, rho = 1)

survdiff(Surv(survival, death) ~ fs.cat + strata(wmsi) + pe, data = ec.data)

```

#### Left-Truncated Right-Censored Survival Data

```{r}
# Transfer the survival time to the age
# ec.data.1 <- ec.data %>% mutate(age = ha.age + round(survival/12)) %>% 
#   filter(age > ha.age)
ec.data <- ec.data %>% mutate(age = ha.age * 12 + survival)
ec.data$age
#fit model
fit.ltrc <- survfit(Surv((ha.age*12), age, death, type = 'counting') ~ 1, data = ec.data)

# Here is a graphical summary of the number of subjects at risk as a function of age (in months).
plot(fit.ltrc$n.risk ~ fit.ltrc$time, type = 'l', lty = 2, col = 'red', 
     main = "Number at Risk as a Function of Age",
     ylab = "Number at Risk", xlab = "Age (Months)", 
     xlim=c(40*12, 1000))

# Condition on living past 55 and 68 years
fit.ltrc1 <- survfit(Surv((ha.age*12), age, death, type = 'counting') ~ 1, data = ec.data, 
                     subset = (ha.age >= 55 ))
fit.ltrc2 <- survfit(Surv((ha.age*12), age, death, type = 'counting') ~ 1, data = ec.data, 
                     subset = (ha.age >= 68 ))
plot(fit.ltrc1$time/12, fit.ltrc1$surv, type = 's', col = 'blue', 
     main = 'Estimated Conditional Survival Functions',
     ylab = 'Estimated Conditional Survival Prob.e', xlab = 'Age (Years)')
lines(fit.ltrc2$time / 12, fit.ltrc2$surv, type = 's', col = 'red')
```


### b) Fitting a Parametric Survival Regression Model
```{r}
#Fitting an intercept-only model (ie. No covariates)
fit.exp <- survreg(Surv(ec.data$survival, ec.data$death) ~ 1, dist = "exponential")
fit.wei <- survreg(Surv(ec.data$survival, ec.data$death) ~ 1, dist = "weibull")
#Results:
summary(fit.exp); summary(fit.wei)



# Adding Covariates
fit.exp1 <- survreg(Surv(ec.data$survival, ec.data$death) ~ ha.age + factor(lvdd.cat) + factor(fs.cat), data = ec.data, dist = "exponential")
summary(fit.exp1)
# NOT significant

# loglogistic
fit.log <- survreg(Surv(ec.data$survival, ec.data$death) ~ ha.age + factor(lvdd.cat) + factor(fs.cat), data = ec.data, dist = "loglogistic")
summary(fit.log)
```

### c) Cox’s Proportional Hazards Model

```{r}

fit <- coxph(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + factor(fs.cat), data = ec.data,
             method='breslow', na.action = na.exclude)
summary(fit)

mg.residual <- resid(fit, type = "martingale")

plot(mg.residual ~ ec.data$ha.age, 
     xlab = "AGE", ylab = "Martingale Residuals",
     main='Martingale Residuals vs. AGE', pch = 19)
#lines(lowess(ec.data$ha.age, mg.residual, delta = 0.5), col = 'red')
```
### Baseline Hazard and Estimation of Survival Function

```{r}
b.haz <- basehaz(fit, centered = F)
t <- b.haz[ ,2]
S.est <- cbind(exp(-b.haz[,1]), t)


### Plot fractional-shortening at 65 yr old
b.est <- fit$coefficients
S0 <- S.est[,1]^(exp(t(b.est) %*% c(65, 0, 0, 0)))
S1 <- S.est[,1]^(exp(t(b.est) %*% c(65, 0, 0, 1)))

plot(S0 ~ t, type = 's', xlim = c(0, 60), ylim = c(0, 1), 
     main = 'Est. Survival Functions For a 65 Year Old \n Compare fractional-shortening', 
     ylab = 'Est. Survival Function', xlab = 'Time')
lines(S1 ~ t, type = 's', lty = 2, col = 2)
legend('bottomleft', 
       c('Abnormal Contracility Age 65', 'Normal Contracility Age 65'),
       col = 1:2, lty = 1:2, bty = 'n', cex = .85)


### wmsi at 65 yr old
b.est <- fit$coefficients
S2 <- S.est[,1]^(exp(t(b.est) %*% c(65, 0, 0, 0)))
S3 <- S.est[,1]^(exp(t(b.est) %*% c(65, 0, 1, 0)))

plot(S2 ~ t, type = 's', xlim = c(0, 60), ylim = c(0, 1), 
     main = 'Est. Survival Functions For a 65 Year Old \n Compare wmsi', 
     ylab = 'Est. Survival Function', xlab = 'Time')
lines(S3 ~ t, type = 's', lty = 2, col = 2)
legend('bottomleft', 
       c('Abnormal Motion Age 65', 'Normal Motion Age 65'),
       col = 1:2, lty = 1:2, bty = 'n', cex = .85)

# 1-year survival probability
res1 <- cbind(S0[max(which(t <= 12))], S1[max(which(t <= 12))])
rownames(res1) <- c('Est. Surv.')
colnames(res1) <- c('Abnormal Conreacility', 'Normal Conreacility')
round(res1, 3)
# 3-year survival probability
res3 <- cbind(S0[max(which(t <= 36))], S1[max(which(t <= 36))])
rownames(res3) <- c('Est. Surv.')
colnames(res3) <- c('Abnormal Conreacility', 'Normal Conreacility')
round(res3, 3)
```

### Time-Dependent Covariates (Section 9.2)

```{r}
ec.data
cut.points <- unique(ec.data$survival[ec.data$death == 1]) #What does this do?
#Converting dataset
ec.data1 <- survSplit(data = ec.data, cut = cut.points, end = "survival", start = "t0", event = "death")
head(ec.data1)

#Create time-dependent covariates
ec.data1$X1 <- ec.data1$pe * log(ec.data1$survival)
ec.data1$X2 <- ec.data1$fs.cat * log(ec.data1$survival)
ec.data1$X3 <- ec.data1$lvdd.cat * log(ec.data1$survival)
ec.data1$X4 <- ec.data1$wmsi * log(ec.data1$survival)

fit.depX1 <- coxph(Surv(t0, survival, death) ~ pe + X1, data = ec.data1, ties = 'breslow')
fit.depX2 <- coxph(Surv(t0, survival, death) ~ fs.cat + X2, data = ec.data1, ties = 'breslow')
fit.depX3 <- coxph(Surv(t0, survival, death) ~ lvdd.cat + X3, data = ec.data1, ties = 'breslow')
fit.depX4 <- coxph(Surv(t0, survival, death) ~ wmsi + X4, data = ec.data1, ties = 'breslow')

fit.ass <- coxph(Surv(t0, survival, death) ~ pe + X1 + fs.cat + X2 + lvdd.cat + X3 + wmsi + X4, data = ec.data1, ties = 'breslow')

```

### Stratified Cox Model

```{r}
# Stratified on PE
fit.pe <- coxph(Surv(t0, survival, death) ~  strata(pe) + ha.age + factor(wmsi) + factor(fs.cat), 
                 data = ec.data1, ties = 'breslow')

ec.pe0 <- ec.data1[ec.data1$pe == 0, ]
ec.pe1 <- ec.data1[ec.data1$pe == 1, ]

fit0 <- coxph(Surv(t0, survival, death) ~  ha.age + factor(wmsi) + factor(fs.cat),
              data = ec.pe0, ties = 'breslow')
fit1 <- coxph(Surv(t0, survival, death) ~  ha.age + factor(wmsi) + factor(fs.cat),
              data = ec.pe1, ties = 'breslow')
X2 <- -2*(fit.pe$loglik[2] - (fit0$loglik[2] + fit1$loglik[2])); X2
1 - pchisq(X2, 3) #3 degrees of freedom for each covariate

#W <- (fit0$coefficients[1] - fit1$coefficients[1])^2 / (fit0$var[1,1] + fit1$var[1,1]); W
#1 - pchisq(W, 1) #p-value


# Stratified on wmsi
fit.wmsi <- coxph(Surv(t0, survival, death) ~  strata(pe) + ha.age + factor(wmsi) + factor(fs.cat), 
                 data = ec.data1, ties = 'breslow')

ec.wmsi0 <- ec.data1[ec.data1$wmsi == 0, ]
ec.wmsi1 <- ec.data1[ec.data1$wmsi == 1, ]

fit0 <- coxph(Surv(t0, survival, death) ~  ha.age + factor(pe) + factor(fs.cat),
              data = ec.wmsi0, ties = 'breslow')
fit1 <- coxph(Surv(t0, survival, death) ~  ha.age + factor(pe) + factor(fs.cat),
              data = ec.wmsi1, ties = 'breslow')
X2 <- -2*(fit.wmsi$loglik[2] - (fit0$loglik[2] + fit1$loglik[2])); X2

1 - pchisq(X2, 3) #3 degrees of freedom for each covariate


# Stratified on fs
fit.fs <- coxph(Surv(t0, survival, death) ~  factor(pe) + ha.age + factor(wmsi) + strata(fs.cat), 
                 data = ec.data1, ties = 'breslow')

ec.fs0 <- ec.data1[ec.data1$fs.cat == 0, ]
ec.fs1 <- ec.data1[ec.data1$fs.cat == 1, ]

fit0 <- coxph(Surv(t0, survival, death) ~  factor(pe) + ha.age + factor(wmsi),
              data = ec.fs0, ties = 'breslow')
fit1 <- coxph(Surv(t0, survival, death) ~  factor(pe) + ha.age + factor(wmsi),
              data = ec.fs1, ties = 'breslow')
X2 <- -2*(fit.fs$loglik[2] - (fit0$loglik[2] + fit1$loglik[2])); X2

1 - pchisq(X2, 3) #3 degrees of freedom for each covariate

```

### d) Graphical Methods to Assess Model Fit for Cox Models

1.  Cox-Snell Residuals

```{r}
#Fit the model
fit <- coxph(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + factor(fs.cat), data = ec.data,
             method='breslow', na.action = na.exclude)
summary(fit)

#Get Cox-Snell residual based on Martingale residuals
mg.residual <- resid(fit, type = "martingale")
cs.residual <- ec.data$death - mg.residual
#Graphical Plot
fit.cs <- survfit(Surv(cs.residual, ec.data$death) ~ 1) #Get Kaplan-Meier estiamtes
H.cs <- cumsum(fit.cs$n.event/fit.cs$n.risk)
plot(fit.cs$time, H.cs, type='s', col='blue', main = 'Cox-Snell Residual Plot')
abline(0, 1, col='red', lty=2)

```
* If the Cox model that we fit was appropriate, then the plot should follow a 45 line. Looking at the plot
above, the model does not fit too badly.

2. Martingale Residuals

```{r}
fit <- coxph(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + factor(fs.cat), data = ec.data,
             method='breslow', na.action = na.exclude)
mg.resid <- resid(fit, type = 'martingale') #Default is Martingale
plot(mg.resid ~ ec.data$ha.age, 
     xlab = "age in years when heart attack occurred ", ylab = "Martingale Residuals",
     main='Martingale Residuals vs. Age', pch = 19)
lines(lowess(ec.data$ha.age, mg.resid, delta = 7), col = 'red')
```

* This is suggestive of ** coding age as an indicator variable **.

3. Graphical Check for Proportional Hazards

* The first method is to stratify on the covariate of interest and plot the log estimated baseline cumulative hazard rates for each strata against time. If the proportionality assumption holds, then the plots should be close to parallel for each strata.


### Parametric Regression Models

```{r}
# Weibull Regression Model
fit.wb <- survreg(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + factor(fs.cat),
                  data = ec.data, dist = "weibull")
summary(fit.wb) #Log-linear Representation
# Fit very well, significant

# Log-Logistic Distribution
fit.ll <- survreg(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + factor(fs.cat), 
                  data = ec.data, dist = "loglogistic")
summary(fit.ll)
# NOT significant 
```


```{r}
source('/Users/huiyuhu/Desktop/Study/UCLA_Biostat/M215/getWeibullEstimates.R')

# Weibull Regression Model
fit.wb <- survreg(Surv(survival, death) ~ ha.age + pe + fs.cat + lvdd.cat + wmsi,
                  data = ec.data, dist = "weibull")
summary(fit.wb) #Log-linear Representation
# Fit very well, significant

sigma <- fit.wb$scale
alpha <- 1 / sigma
eta <- -fit.wb$linear.predictors / sigma

r.wb <- ec.data$survival^alpha * exp(eta)

fit.wb1 <- survfit(Surv(r.wb, ec.data$death) ~ 1)
H.wb <- cumsum(fit.wb1$n.event/fit.wb1$n.risk)

plot(H.wb ~ fit.wb1$time, type = 'l', main = 'Cox-Snell Residual Plot for \n Weibull Regression',
ylab = 'Estimated Cumulative Hazard', xlab = 'Cox-Snell Residual')
abline(0, 1, col='red', lty=2)



getWeibullEstimates(fit.wb)

# Log-Logistic Distribution
fit.ll <- survreg(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + factor(fs.cat), 
                  data = ec.data, dist = "loglogistic")
summary(fit.ll)
# NOT significant 
```

```{r}
table(ec.data$lvdd.cat)
table(ec.data$fs.cat)
table(ec.data$wmsi)
# fit.p <- coxph(Surv(survival, death) ~ ha.age + factor(pe) + factor(wmsi) + 
#                  epss + factor(lvdd.cat)+ epss*factor(lvdd.cat), 
#                data = ec.data2,method='breslow', na.action = na.exclude)
# summary(fit.p)


```

```{r}
# univariate test and full model

covariates <- c("ha.age", "fs.cat",  "wmsi", "lvdd.cat", "pe")
univ_formulas <- sapply(covariates,
                        function(x) as.formula(paste('Surv(survival, death)~', x)))
                        
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = ec.data)})
# Extract data 
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = FALSE))
as.data.frame(res)

res.cox <- coxph(Surv(survival, death) ~ ha.age+ fs.cat + wmsi + lvdd.cat + pe , data = ec.data)
summary(res.cox)
```


















