---
title: "M215_HW6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(survival) # loading survival functions into R
library(KMsurv) # datasets from Klein and Moeschberger textbook
help(btrial)
```

## 8.4
a.
```{r}
help(hodg)
data(hodg)
hodg$Z1 <- ifelse(hodg$gtype == 1 & hodg$dtype == 2, 1, 0)
hodg$Z2 <- ifelse(hodg$gtype == 2 & hodg$dtype == 1, 1, 0)
hodg$Z3 <- ifelse(hodg$gtype == 2 & hodg$dtype == 2, 1, 0)
fit <- coxph(Surv(time, delta) ~ Z1+Z2+Z3,ties = 'breslow', data = hodg)
summary(fit)
```

b. 
```{r}
hodg$Z_1 <- ifelse(hodg$gtype == 2, 1, 0)
hodg$Z_2<- ifelse(hodg$dtype == 2, 1, 0)
hodg$Z12 <- hodg$Z_1 * hodg$Z_2
fit1 <- coxph(Surv(time, delta) ~ Z_1+Z_2+Z12,ties = 'breslow', data = hodg)
summary(fit1)
```
(c) Find point estimates and 95% confidence intervals for the relative risk of death for an NHL Auto transplant patient as compared to an NHL Allo transplant patient.

**In the summary of the first model, the relative risk of death fro an NHL Auto transplant patients as compared to an NHL Allo transplant patient is the exponential coefficient of of Z2a which is 1.94 in this case with the 95% CI of (0.64,5.87).**

(d) Find the p-value of a test of the hypothesis that the hazard rates are the same for HOD Allo transplants and NHL Allo patients, using the Wald test. Repeat a similar test for Auto patients.

```{r}
#Allo patient
C <- c(1, 0, 0)
b0 <- c(0, 0, 0)
b <- fit.1$coefficients
V <- fit.hodg$var
chi_allo <- t(C %*% b - C %*% b0) %*% solve(t(C) %*% V %*% C) %*% (C %*% b - C %*% b0)
pchisq(chi_allo, df=1, lower.tail = F)

#Auto patient
C <- c(0, 1, -1)
b0 <- c(0, 0, 0)
b <- fit.hodg$coefficients
V <- fit.hodg$var
chi_allo <- t(C %*% b - C %*% b0) %*% solve(t(C) %*% V %*% C) %*% (C %*% b - C %*% b0)
pchisq(chi_allo, df=1, lower.tail = F)
```

(e) Test the hypothesis, using the Wald test, that the hazard rates for Auto transplant and Allo transplant patients are the same for each disease group against the alternative that the hazard rates for Auto transplant and Allo transplant patients for at least one group are dif- ferent using a two-degree of freedom test of H0 : h(t | NHL Allo) =􏰗 h(t |NHL Auto) and H0 : h(t |HOD Allo) =􏰗 h(t |HOD Auto)

```{r}
C <- rbind(c(0, 1, 0), c(1, 0, -1))
b0 <- c(0, 0, 0)
b <- fit.hodg$coefficients
V <- fit.hodg$var
chi_allo <- t(C %*% b - C %*% b0) %*% solve(C %*% V %*% t(C)) %*%
(C %*% b - C %*% b0)
pchisq(chi_allo, df=2, lower.tail = F)
```



## 8.10

a)
```{r}
help(bmt)
data(bmt)
coxph(Surv(ta, da) ~ z10, data = bmt, ties = 'breslow')
```
b) 
```{r}
bmt$Z1 <- ifelse(bmt$group == 1, 1, 0)
bmt$Z2 <- ifelse(bmt$group == 2, 1, 0)
bmt$Z1_MTX <- bmt$Z1 * bmt$z10
bmt$Z2_MTX <- bmt$Z2 * bmt$z10
coxph(Surv(ta, da) ~ z10 + Z1 + Z2 + Z1_MTX + Z2_MTX, data = bmt, ties = 'breslow')

```
c) 

d) 
```{r}
bmt$Q1 <- ifelse(bmt$group == 2, 1, 0)
bmt$Q2 <- ifelse(bmt$group == 3, 1, 0)
bmt$Q3 <- bmt$z7
bmt$Q4 <- ifelse(bmt$z8 == 1, 1, 0)
bmt$Q5 <- bmt$z10
bmt$Q6 <- ifelse(bmt$z4 == 1, 1, 0) 
bmt$Q7 <- ifelse(bmt$z3 == 1, 1, 0)
bmt$Q8 <- bmt$Q6 * bmt$Q7
bmt$Q9 <- ifelse(bmt$z6 == 1, 1, 0)
bmt$Q10 <-ifelse(bmt$z5 == 1, 1, 0)
bmt$Q11 <- bmt$Q9 * bmt$Q10
bmt$Q12 <- bmt$z2 - 28
bmt$Q13 <- bmt$z1 - 28
bmt$Q14 <- bmt$Q12 * bmt$Q13

initial.model <- coxph(Surv(ta, da) ~ Q5, data = bmt, ties = "breslow")

library(MASS)
stepAIC(initial.model, direction = "both", scope = list(upper = ~ Q1 + Q2 + Q3 +Q4+Q5+Q6+Q7+Q8+Q9+Q10+Q11+Q12+Q13+Q14, lower = ~ Q5))
```

## 8.14
a) Estimate the survival functions of the time to AGVHD for the MTX and no MTX treatment groups discussed in Exercise 8.10, adjusted for disease category. Provide a separate estimate for each disease group.

```{r}
bmt$Z1 <- ifelse(bmt$group == 1, 1, 0)
bmt$Z2 <- ifelse(bmt$group == 2, 1, 0)
bmt$Z1_MTX <- bmt$Z1 * bmt$z10
bmt$Z2_MTX <- bmt$Z2 * bmt$z10
#according to 8.10, interaction effects are not sinificant 
fit.bmt2 <- coxph(Surv(ta, da) ~ z10 + factor(group), data = bmt, ties = 'breslow')
fit.bmt2

b.haz <- basehaz(fit.bmt2, centered = F)
t <- b.haz[ ,2]
S.est <- cbind(exp(-b.haz[,1]), t)
b.est <- fit.bmt2$coefficients
S1_n<- S.est[,1]^(exp(t(b.est) %*% c(0, 0, 0))) # group 1 no MTX
S2_n <- S.est[,1]^(exp(t(b.est) %*% c(1, 0, 0))) # group 2 no MTX
S3_n <- S.est[,1]^(exp(t(b.est) %*% c(0, 1, 0))) # group 3 no MTX

S1 <- S.est[,1]^(exp(t(b.est) %*% c(0, 0, 1))) # group 1 
S2 <- S.est[,1]^(exp(t(b.est) %*% c(1, 0, 1))) # group 2
S3 <- S.est[,1]^(exp(t(b.est) %*% c(0, 1, 1))) # group 3

res <- cbind(S1_n, S2_n, S3_n, S1, S2, S3)
colnames(res) <- c("ALL-no MTX", "AML low-no MTX", "AML high-no MTX",
                   "ALL- MTX", "AML low- MTX", "AML high- MTX")
res <- cbind(S1[max(which(t <= 80))], S2[max(which(t <= 80))],
S3[max(which(t <= 80))], S4[max(which(t <= 80))])
rownames(res) <- c('Est. Surv.')
colnames(res) <- c('base', 'MTX', 'ALL', 'AML HIGH')
round(res, 3)

fit.s1 <- survfit(fit.bmt2, newdata = data.frame(group = 3, z10 = 0),
se.fit = TRUE, conf.int = .95, conf.type = 'log-log')
summary(fit.s1)

fit.s2 <- survfit(fit.bmt2, newdata = data.frame(group = 3, z10 = 1),
se.fit = TRUE, conf.int = .95, conf.type = 'log-log')
summary(fit.s2)
```




## 9.1
```{r}
help("tongue")
data(tongue)
cut.points <- unique(tongue$time[tongue$delta == 1]) 
tongue1 <- survSplit(data = tongue, cut = cut.points, end = "time", start = "t0", event = "delta")

tongue1$X1 <- ifelse(tongue1$type == 2 , 1, 0)
tongue1$X2 <- tongue1$X1 * log(tongue1$time)
fit.tongue <- coxph(Surv(t0, time, delta) ~ X1 + X2, data = tongue1, ties = 'breslow')
summary(fit.tongue)
```


## 9.3 

```{r}
time1 <- c(1, 63, 105, 129, 182, 216, 250, 262, 301, 301, 342, 354, 356, 358, 380, 383, 383, 388, 394, 408, 460, 489, 499, 523, 524, 535, 562, 569, 675, 676, 748, 778, 786, 797, 955, 968, 1000, 1245, 1271, 1420, 1551, 1694, 2363, 2754, 2950)

delta1 <- rep(c(1, 0), times = c(43, 2))

# group == 1 is chemotherapy only
group1 <- rep(1, 45)


time2 <- c(17, 42, 44, 48, 60, 72, 74, 95, 103, 108, 122, 144, 167, 170, 183, 185, 193, 195, 197, 208, 234, 235, 254, 307, 315, 401, 445, 464, 484, 528, 542, 547, 577, 580, 795, 855, 1366, 1577, 2060, 2412, 2486, 2796, 2802, 2934, 2988)


delta2 <- rep(c(1, 0), times = c(39, 6))
# group == 0 is chemotherapy plus radiotherapy
group2 <- rep(0, 45)

data <- cbind(c(time1, time2), c(delta1, delta2), c(group1, group2))
colnames(data) <- c("time", "delta","group")

data <- as.data.frame(data)
```
```{r}
fit.ct <- coxph(Surv(time, delta) ~ group, data = data, ties = "breslow")

summary(fit.ct)
```

```{r}
cut.points_non <- unique(data$time[data$delta == 1]) 
data1 <- survSplit(data = data, cut = cut.points_non, end = "time", start = "t0", event = "delta")
data1$X1 <- data1$group * log(data1$time)
fit.non <- coxph(Surv(t0, time, delta) ~ group + X1, data = data1, ties = 'breslow')
summary(fit.non)
```

```{r}
loglik <- as.numeric(length(cut.points_non))
for(i in 1:length(cut.points_non)){
data1$tdc1 <- ifelse(data1$time > cut.points_non[i], data1$group, 0)
fit.tdc <- coxph(Surv(t0, time, delta) ~ group + tdc1, data = data1, ties = 'breslow')
loglik[i] <- fit.tdc$loglik[2]
}
cbind(cut.points_non, loglik)
opt_tau <- cut.points_non[which.max(loglik)]
opt_tau
```

```{r}
data1$Z2 <- ifelse(data1$time > opt_tau, data1$group, 0)
data1$Z3 <- ifelse(data1$time <= opt_tau, data1$group, 0)
fit.opt <- coxph(Surv(t0, time, delta) ~ Z2 + Z3, data = data1, ties = 'breslow')
summary(fit.opt)
```




## 9.5
```{r}
data(larynx)
help(larynx)
#Create dummy variables for stage variable (s2, s3, s4)
larynx$s2 <- ifelse(larynx$stage == 2, 1, 0)
larynx$s3 <- ifelse(larynx$stage == 3, 1, 0)
larynx$s4 <- ifelse(larynx$stage == 4, 1, 0)
larynx$yr <- ifelse(larynx$diagyr < 75, 1, 0)

cut.points <- unique(larynx$time[larynx$delta == 1]) 
larynx1 <- survSplit(data = larynx, cut = cut.points, end = "time", start = "t0", event = "delta")
# stratify based on year
fit <- coxph(Surv(t0, time, delta) ~ s2+s3+s4+age+strata(yr), data = larynx1, ties = 'breslow')
fit
```

b)
```{r}
larynx_yr0 <- larynx1[larynx1$yr == 0, ]
larynx_yr1 <- larynx1[larynx1$yr == 1, ]
fit0 <- coxph(Surv(t0, time, delta) ~ s2+s3+s4+age, data = larynx_yr0, ties = 'breslow')

fit1 <- coxph(Surv(t0, time, delta) ~ s2+s3+s4+age, data = larynx_yr1, ties = 'breslow')

#LR
X2 <- -2*(fit$loglik[2] - (fit0$loglik[2] + fit1$loglik[2])); X2
1 - pchisq(X2, 4)

#WALT
W1 <- (fit0$coefficients[1] - fit1$coefficients[1])^2 / (fit1$var[1,1] + fit0$var[1,1]); W1
1 - pchisq(W1, 1) #p-value

W2 <- (fit0$coefficients[2] - fit1$coefficients[2])^2 / (fit1$var[2,2] + fit0$var[2,2]); W2
1 - pchisq(W2, 1) #p-value

W3 <- (fit0$coefficients[3] - fit1$coefficients[3])^2 / (fit1$var[3,3] + fit0$var[3,3]); W3
1 - pchisq(W3, 1) #p-value

W4 <- (fit0$coefficients[4] - fit1$coefficients[4])^2 / (fit1$var[4,4] + fit0$var[4,4]); W4
1 - pchisq(W4, 1) #p-value

C <-rbind(c(1, 0, 0, 0), c(0, 1, 0, 0), c(0, 0, 1, 0), c(0, 0, 0, 1))
b0 <-fit0$coefficients
b1 <- fit1$coefficients
V <- fit0$var + fit1$var
wald <- t(b1 - b0) %*% solve(C %*% V %*% t(C)) %*% (b1 - b0)

wald
1- pchisq(wald, 4)

```


## 9.8

```{r}
help(burn)
data(burn)

cut.points <- unique(burn$T3[burn$D3 == 1])

burn1 <- survSplit(data = burn, cut = cut.points, end = "T3", start = "t0", event = "D3")

#Create time-dependent covariates
burn1$co <- ifelse(burn1$T3 >= burn1$T1 & burn1$D1 == 1, 1, 0)

fit <- coxph(Surv(t0, T3, D3) ~ co, data = burn1, ties = 'breslow')
summary(fit)
```


b) 
```{r}
burn1$co1 <- ifelse(burn1$T3 >= burn1$T2 & burn1$D2 == 1, 1, 0)

fit1 <- coxph(Surv(t0, T3, D3) ~ co1, data = burn1, ties = 'breslow')

summary(fit1)
```

```{r}
burn1$Z1_logt <- burn1$Z1 * log(burn1$T3)
burn1$Z2_logt <- burn1$Z2 * log(burn1$T3)
burn1$Z3_logt <- burn1$Z3 * log(burn1$T3)
burn1$Z4_logt <- burn1$Z4 * log(burn1$T3)
burn1$Z5_logt <- burn1$Z5 * log(burn1$T3)
burn1$Z6_logt <- burn1$Z6 * log(burn1$T3)
burn1$Z7_logt <- burn1$Z7 * log(burn1$T3)
burn1$Z8_logt <- burn1$Z8 * log(burn1$T3)
burn1$Z9_logt <- burn1$Z9 * log(burn1$T3)
burn1$Z10_logt <- burn1$Z10 * log(burn1$T3)
burn1$co_logt <- burn1$co * log(burn1$T3)
burn1$co1_logt <- burn1$co1 * log(burn1$T3)
burn1$Z11_1 <- ifelse(burn1$Z11 == 2, 1, 0)
burn1$Z11_2 <- ifelse(burn1$Z11 == 3, 1, 0)
burn1$Z11_3 <- ifelse(burn1$Z11 == 4, 1, 0)
burn1$Z11_1_logt <- burn1$Z11_1 * log(burn1$T3)
burn1$Z11_2_logt <- burn1$Z11_2 * log(burn1$T3)
burn1$Z11_3_logt <- burn1$Z11_3 * log(burn1$T3)
fit.burn_1 <- coxph(Surv(t0, T3, D3) ~  Z1+Z2+Z3+Z4+Z5+Z6+Z7+Z8+Z9+Z10+factor(Z11)+co+co1+Z9_logt, data = burn1, ties = 'breslow')
summary(fit.burn_1)
fit.burn_2 <- coxph(Surv(t0, T3, D3) ~  Z1+Z2+Z3+Z4+Z5+Z6+Z7+Z8+Z9+Z10+factor(Z11)+co+co1+Z11_1_logt, data = burn1, ties = 'breslow')
summary(fit.burn_2)
fit.burn_3 <- coxph(Surv(t0, T3, D3) ~  Z1+Z2+Z3+Z4+Z5+Z6+Z7+Z8+Z9+Z10+factor(Z11)+co+co1+Z11_3_logt, data = burn1, ties = 'breslow')
summary(fit.burn_3)

```












